Giant Low Surface Brightness Galaxies As Merger Remnants

### 2015-04-24 glsb.tex
blah
working on intro

\altaffiltext{1}{Department of Astronomy, University of Washington, Box 351580,
Seattle WA, 98195}
\altaffiltext{2}{XXX-U Mich}
\altaffiltext{3}{Jeremiah Horrocks Institute, University of Central Lancashire, Preston PR1 2HE, UK.  RCUK Fellow}
\altaffiltext{4}{XXX-cal tech}

@@

Since the discovery of Malin~1 \citep{Bothun87}, giant low surface brightness galaxies (GLSB) have presented a unique problem for galaxy formation theories.  Low surface brightness (LSB) galaxies are typically defined as disks with central surface brightnesses fainter than 23 $B$ mag arcsec$^2$.  These disks are incredibly large, with Malin~1 being detected to a radius of over 100 kpc \citep{Moore06}.  It is difficult to imagine how a stellar disk could form within a merger-driven $\Lambda$CDM cosmology.

A wide variety of formation scenarios have been proposed for GLSB systems.  \citet{Mapelli08} form GLSB-like galaxies via the fading of collisional ring galaxies xxx-expand this a bit.

Meanwhile, observers tend to classify GLSB systems as normal spirals.  \citet{Matthews01b} list Malin~2 as an Scd while \citet{Das13} concludes that GLSB systems are an extreme form of late type spiral galaxies.  Similarly, \citet{Kasparova14} claim that the large dark matter scale of Malin~2 can explain it's features.

@@

Standard catalogs list these systems as Sd and Sa galaxies, so they are commonly thought of as spirals.

@@

Outline for discussion section
\begin{itemize}
\item{We find the GLSB systems Malin~2 and UGC~6614 are observationally most consistent with being normal elliptical galaxies that have undergone a high angular momentum minor merger with a gas rich system.}
\item{recap photometric evidence--the centers look just like SDSS ellipticals.}
\item{recap star formation histories. Disk is dominated by old stars.}
\item{Discuss metallicity gradient. Hard to imagine how such a metal poor disk could form right next to a very metal rich core.}
\item{extended comparison to \citet{Kasparova14}}
  \begin{itemize}
    \item{Mass model of Malin~2 requires a huge isothermal core radius.}
    \item{They measure gas metallicities and find solar in the center and half-solar at 20-30 kpc.  While it is a fool's errand to compare metallicities between stars and gas, it is interesting that the gas metallicity gradient is much smaller than in the stars.  If the central stellar metallicity is indeed super-solar and the gas metallicity only solar, that implies there must have been low-metallicity gas accretion onto the system at some point.  Hmm, looking at DR12 stuff, the pipeline reducitons are fitting solar metallicity to the center, not sure I believe that.  }
      \item{They have a nice mass-model that becomes dark matter dominated at around r=20 kpc.  Could this be the normal shape of elliptical dark matter halos?}
  \end{itemize}

>>>>

\altaffiltext{1}{Department of Astronomy, University of Washington, Box 351580,
Seattle WA, 98195}
\altaffiltext{2}{XXX-U Mich}
\altaffiltext{3}{Jeremiah Horrocks Institute, University of Central Lancashire, Preston PR1 2HE, UK.  RCUK Fellow}
\altaffiltext{4}{XXX-Cal Tech}

@@

Since the discovery of Malin~1 \citep{Bothun87}, giant low surface brightness galaxies (GLSB) have presented a unique problem for galaxy formation theories.  Low surface brightness (LSB) galaxies are typically defined as disks with central surface brightnesses fainter than 23 $B$ mag arcsec$^2$.  While LSB galaxies span a wide range of sizes and masses, the few giant low surface brightness that have been discovered seem to be unrelated objects that are at least an order of magnitude larger in extent than typical LSB systems (XXX-cite).

Observers tend to classify GLSB systems as normal spirals.  \citet{Matthews01b} list Malin~2 as an Scd while \citet{Das13} concludes that GLSB systems are an extreme form of late type spiral galaxies.  Similarly, \citet{Kasparova14} claim that the large dark matter scale of Malin~2 can explain it's features.  Even the Galaxy Zoo project classifies GLSB galaxies as spirals.

The disks in GLSB systems are incredibly large, with Malin~1 being detected to a radius of over 100 kpc \citep{Moore06}.  Disks of this size are a challenge for typicall $\Lambda$CDM cosmology galaxy formation and exotic formation scenarios have been proposed for GLSB systems.  \citet{Mapelli08} form GLSB-like galaxies via the fading of collisional ring galaxies xxx-expand this a bit.  In their model, a collision where a thingy hits thuswise, blah blah, sets off an expanding ring of star formation.  As this the star formation subsides and the stetllar population ages, it fades to form a disk that resembles in size and exponential profile observed GLSBs.  XXX--also tie in \citep{Mapelli12}.

@@

@@

Outline for discussion section
\begin{itemize}
\item{We find the GLSB systems Malin~2 and UGC~6614 are observationally most consistent with being normal elliptical galaxies that have undergone a high angular momentum minor merger with a gas rich system.}
\item{recap photometric evidence--the centers look just like SDSS ellipticals.}
\item{recap star formation histories. Disk is dominated by old stars.}
\item{Discuss metallicity gradient. Hard to imagine how such a metal poor disk could form right next to a very metal rich core.
      \citet{Raskutti14} use VIRUS-P to look at stellar halo kinematics of massive ellipticals.  Lots of good stuff on fast/slow rotators.  Also a very nice plot showing the radial gradients in age and metallicity as measured from Lick indices.  They don't see any sharp transition like I do, all very smooth age and metal curves. }
\item{extended comparison to \citet{Kasparova14}}
  \begin{itemize}
    \item{Mass model of Malin~2 requires a huge isothermal core radius.}
    \item{They measure gas metallicities and find solar in the center and half-solar at 20-30 kpc.  While it is a fool's errand to compare metallicities between stars and gas, it is interesting that the gas metallicity gradient is much smaller than in the stars.  If the central stellar metallicity is indeed super-solar and the gas metallicity only solar, that implies there must have been low-metallicity gas accretion onto the system at some point.  Hmm, looking at DR12 stuff, the pipeline reducitons are fitting solar metallicity to the center, not sure I believe that.  }
      \item{They have a nice mass-model that becomes dark matter dominated at around r=20 kpc.  Could this be the normal shape of elliptical dark matter halos?}
  \end{itemize}
